---
title: "Week 5 Video Script"
format: html
editor: visual
---

## EDA principles

1. Generate questions about your data. 

- __What impacts the number of bike rentals?__

2. Search for answers by visualizing, transforming, and modelling your data.

3. Use what you learn to refine your questions and/or generate new questions.



## ggplot Layers

```{r, message = F}
library(tidyverse)
bike_counts <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/Bike.csv')

glimpse(bike_counts)
head(bike_counts)
summary(bike_counts)
```

---

#### 1D: explore count

```{r}
bike_counts |>
  ggplot(aes(x= count)) +
  geom_histogram(bins = 100, color = 'blue', fill = 'yellow') +
  theme_bw()
```


#### 2D: explore count vs. temperature

```{r}
bike_counts |>
  mutate(season_factor = factor(season),
         holiday_factor = factor(holiday)) |>
  ggplot(aes(y = count, x = temp, color = season_factor, shape = holiday_factor)) +
  geom_jitter(alpha = .1) +
  geom_smooth(method = 'loess', formula = 'y ~ x') +
  theme_bw() +
  facet_wrap(season_factor~.)
```

#### 2D: explore count vs. season (factor)

```{r}
bike_counts |>
  mutate(season_factor = factor(season),
         holiday_factor = factor(holiday)) |>
  ggplot(aes(y = count, x = season_factor, shape = holiday_factor, color = holiday_factor)) +
  geom_violin() +
  geom_boxplot() +
    geom_jitter(alpha = .1) +
  theme_bw() +
  facet_wrap(holiday_factor ~ .)
```


---

### Aesthetics

-   Color
-   Shape
-   Size
-   x-axis
-   y-axis

---

### Geometric Objects

-   geom_point() / geom_jitter()
-   geom_line() / geom_smooth()
-   geom_boxplot()
-   geom_histogram()

### Facets
